---
title: "elastic net"
output:
  pdf_document:
    latex_engine: pdflatex  # or try xelatex or lualatex
header-includes:
  - \usepackage{booktabs} 
date: "2024-11-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
two dataset:
"final_full_rmna_OneHotEncoded_housing.csv" (no outliers removed)
"final_rmna_OneHotEncoded_housing.csv" (price outliers removed)

# (1)dataset with all complete cases (no missing values)
```{r}
library(readr)
library(glmnet)

housing_data <- read_csv("/Users/chensihan/Documents/GitHub/BIOSTAT625/Data Exploration/final_full_rmna_OneHotEncoded_housing.csv")
set.seed(34)
```

```{r}
attach(housing_data)
#transform price into log_price
housing_data$log_price=log(housing_data$`Price (in rupees)`)
X <- housing_data[, -which(names(housing_data) == "Price (in rupees)" | names(housing_data) == "log_price")] 
Y <- housing_data$log_price
X_matrix <- as.matrix(X)
Y_matrix <- as.matrix(Y)
```

```{r}
# Initialize storage
alphas <- seq(0, 1, by = 0.01)
elastic_net <- list()  # A list to store models
best_lambda <- numeric(length(alphas))  # Numeric vector for best lambda values
predicted_prices <- list()  # A list to store predictions
results <- list()  # A list to store data frames with actual and predicted values
rmse <- numeric(length(alphas))   # Numeric vector for RMSE values
mae <- numeric(length(alphas))   # Numeric vector for MAE values
mape=numeric(length(alphas))  # Numeric vector for MAPE values

for (i in seq_along(alphas)) {
  alpha=alphas[i]
  # Fit elastic net model with alpha
  elastic_net[[i]] <- cv.glmnet(X_matrix, Y_matrix, alpha = alpha)
  
  # Store the best lambda value (we choose lambda where MSE is smallest) #lambda.min
  best_lambda[i] <- elastic_net[[i]]$lambda.1se
  
  # Predict on the test set using the best lambda
  predicted_prices[[i]] <- predict(elastic_net[[i]], newx = X_matrix, s = "lambda.1se")
  
  # Combine actual and predicted values
  results[[i]] <- data.frame(Actual = Y_matrix, Predicted = as.vector(predicted_prices[[i]]))
  # Compute RMSE, MAE, MAPE
  rmse[i] <- sqrt(mean((results[[i]]$Actual - results[[i]]$Predicted)^2))
  mae[i] <- mean(abs(results[[i]]$Actual - results[[i]]$Predicted))
  mape[i] <- mean(abs((results[[i]]$Actual - results[[i]]$Predicted) / results[[i]]$Actual)) * 100
}
```

```{r}
# Results for RMSE, MAE and MAPE
print(rmse)
which.min(abs(rmse))
```

```{r}
print(mae)
which.min(abs(mae))
```

```{r}
print(mape)
which.min(abs(mape))
```
## best alpha is 0.05
```{r}
plot(alphas,rmse)
plot(alphas,mape)
plot(alphas, mae)

alphas[6]
```

## we choose lambda.1se = 0.08663284
```{r}
fit= cv.glmnet(X_matrix, Y_matrix, alpha = alphas[6])
plot(fit)
# two methods to choose lambda
fit$lambda.1se
fit$lambda.min
```

s comes from lambda.1se

```{r}
# Compute RMSE, MAE, MAPE and R_squared in training dataset
predicted_prices<- predict(fit, newx = X_matrix,s=0.08663284)
  
  # Combine actual and predicted values
results<- data.frame(Actual = Y_matrix, Predicted = as.vector(predicted_prices))
  
  rmse<- sqrt(mean((results$Actual - results$Predicted)^2))
  mae<- mean(abs(results$Actual - results$Predicted))
  mape<- mean(abs((results$Actual - results$Predicted) / results$Actual)) * 100
  rmse
  mae
  mape

plot(Y_matrix, predicted_prices, xlab = "Actual Prices (natural log)", ylab = "Predicted Prices (natural log)" )

y_mean <- mean(Y_matrix)
SS_total <- sum((Y_matrix - y_mean)^2)
SS_residual <- sum((Y_matrix - predicted_prices)^2)
R_squared <- 1 - (SS_residual / SS_total)
print(R_squared)

```
```{r}
# Determine the range of both axes
range_values <- range(c(Y_matrix, predicted_prices))

# Create the plot with the same scale for x and y axes
plot(
  Y_matrix, predicted_prices,
  xlab = "Actual Prices (natural log)",
  ylab = "Predicted Prices (natural log)",
  xlim = range_values, 
  ylim = range_values,
  main = "Actual vs Predicted Prices in Dataset(1)"
)

# Optionally, add a 45-degree line for reference
abline(0, 1, col = "red", lty = 2)  # A red dashed line

```
# (2)a refined dataset excluding both missing values and outliers
```{r}
housing_data2 <- read_csv("/Users/chensihan/Documents/GitHub/BIOSTAT625/Data Exploration/final_rmna_OneHotEncoded_housing.csv")
attach(housing_data2)

#transform price into log_price
housing_data2$log_price=log(housing_data2$`Price (in rupees)`)
X2 <- housing_data2[, -which(names(housing_data2) == "Price (in rupees)" | names(housing_data2) == "log_price")] 
Y2 <- housing_data2$log_price
X2_matrix <- as.matrix(X2)
Y2_matrix <- as.matrix(Y2)
```


```{r}
# Initialize storage
alphas2 <- seq(0, 1, by = 0.01)
elastic_net2 <- list()  # A list to store models
best_lambda2 <- numeric(length(alphas2))  # Numeric vector for best lambda values
predicted_prices2 <- list()  # A list to store predictions
results2 <- list()  # A list to store data frames with actual and predicted values
rmse2 <- numeric(length(alphas2))   # Numeric vector for RMSE values
mae2 <- numeric(length(alphas2))   # Numeric vector for MAE values
mape2=numeric(length(alphas2))  # Numeric vector for MAPE values

for (i in seq_along(alphas2)) {
  alpha2=alphas2[i]

  elastic_net2[[i]] <- cv.glmnet(X2_matrix, Y2_matrix, alpha = alpha2)
  
  # Store the best lambda value (we choose lambda where MSE is smallest) 
  best_lambda2[i] <- elastic_net2[[i]]$lambda.1se
  
  # Predict on the test set using the best lambda
  predicted_prices2[[i]] <- predict(elastic_net2[[i]], newx = X2_matrix, s = "lambda.1se")
  
  # Combine actual and predicted values
  results2[[i]] <- data.frame(Actual = Y2_matrix, Predicted = as.vector(predicted_prices2[[i]]))
  # Compute RMSE, MAE, MAPE
  rmse2[i] <- sqrt(mean((results2[[i]]$Actual - results2[[i]]$Predicted)^2))
  mae2[i] <- mean(abs(results2[[i]]$Actual - results2[[i]]$Predicted))
  mape2[i] <- mean(abs((results2[[i]]$Actual - results2[[i]]$Predicted) / results2[[i]]$Actual)) * 100
}
```

```{r}
# Results for RMSE, MAE and MAPE
print(rmse2)
which.min(abs(rmse2))
```

```{r}
print(mae2)
which.min(abs(mae2))
```

```{r}
print(mape2)
which.min(abs(mape2))
```

```{r}
plot(alphas2,rmse2)
plot(alphas2,mape2)
plot(alphas2, mae2)
```
## best alpha is 0.05
```{r}
fit2= cv.glmnet(X2_matrix, Y2_matrix, alpha = 0.05)
plot(fit2)
# two methods to choose lambda
fit2$lambda.1se
fit2$lambda.min
```
## we choose lambda.1se = 0.0342591
s comes from lambda.1se
```{r}
# Compute RMSE, MAE, MAPE and R_squared in training dataset
predicted_prices2<- predict(fit2, newx = X2_matrix,s=0.0342591)
  
  # Combine actual and predicted values
results2<- data.frame(Actual = Y2_matrix, Predicted = as.vector(predicted_prices2))
  
  rmse2<- sqrt(mean((results2$Actual - results2$Predicted)^2))
  mae2<- mean(abs(results2$Actual - results2$Predicted))
  mape2<- mean(abs((results2$Actual - results2$Predicted) / results2$Actual)) * 100
  rmse2
  mae2
  mape2

y2_mean <- mean(Y2_matrix)
SS_total2 <- sum((Y2_matrix - y2_mean)^2)
SS_residual2 <- sum((Y2_matrix - predicted_prices2)^2)
R_squared2 <- 1 - (SS_residual2 / SS_total2)
print(R_squared2)

```


```{r}
# Determine the range of both axes
range_values <- range(c(Y2_matrix, predicted_prices2))

# Create the plot with the same scale for x and y axes
plot(
  Y2_matrix, predicted_prices2,
  xlab = "Actual Prices (natural log)",
  ylab = "Predicted Prices (natural log)",
  xlim = range_values, 
  ylim = range_values,
  main = "Actual vs Predicted Prices in Dataset(2)"
)

# Optionally, add a 45-degree line for reference
abline(0, 1, col = "red", lty = 2)  # A red dashed line

```
# Actual vs Predicted Prices Plot
```{r}
# Set layout to display 2 plots side by side
par(mfrow = c(1, 2))  # 1 row, 2 columns

# Plot 1
range_values <- range(c(Y_matrix, predicted_prices))
plot(
  Y_matrix, predicted_prices,
  xlab = "Actual Prices (natural log)",
  ylab = "Predicted Prices (natural log)",
  xlim = range_values, 
  ylim = range_values,
  cex.lab = 0.8, 
  cex.main = 0.8,  
  main = "Actual vs Predicted Prices in Dataset(1)"
)
abline(0, 1, col = "red", lty = 2)  # A red dashed line

# Plot 2

plot(
  Y2_matrix, predicted_prices2,
  xlab = "Actual Prices (natural log)",
  ylab = "Predicted Prices (natural log)",
  xlim = range_values, 
  ylim = range_values,
  cex.lab = 0.8, 
  cex.main = 0.8,  
  main = "Actual vs Predicted Prices in Dataset(2)"
)
abline(0, 1, col = "red", lty = 2)  # A red dashed line

# Reset layout to default (1 plot)
par(mfrow = c(1, 1))

```
```{r}
library(knitr)

# Create a data frame
performance_metrics <- data.frame(
  `Performance Metric` = c("RMSE", "MAE", "MAPE", "R-square"),
  `Dataset$(1)$` = c(0.390, 0.290, 3.279, 0.447),
  `Dataset$(2)$` = c(0.382, 0.283, 3.211, 0.463)
)

# Display the table using knitr::kable with booktabs formatting
kable(performance_metrics, format = "latex", booktabs = TRUE, caption = "Model Performance Metrics for Dataset 1 and Dataset 2")

```

