---
title: "HousePrice_descriptive"
author: "GuoJennifer"
date: "2024-10-30"
output: html_document
---

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 70)
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Download and Summary

Get data set.

```{r mysize=T,size='\\footnotesize', echo=T,eval=T}

setwd("~/Desktop/625 final project") 

#library(psych)
library(ggplot2)
library("corrplot")
library("gtsummary")
#library("dplyr")

data = read.csv(file="house_prices.csv",header=T)
attach(data)
head(data)
#summary(data)

df = data.frame(data)
```

## Cleaning and Encoding

Cleaning data and one-hot encoding for Furnishing/ Facing/ Overlooking

## Furnishing

Levels for Furnishing are either Unfurnished, Semi-Furnished, or Furnished
Subjects with missing data are indicated with "1" in the last column of the one-hot matrix.

```{r mysize=T,size='\\footnotesize', echo=T,eval=T}
#summary(data$Furnishing)
#psych::describe(data$Furnishing)
data$Furnishing = as.factor(data$Furnishing)
unique(data$Furnishing)
ggplot(data, aes(x=Furnishing)) + geom_bar()

onehot_Furnishing = model.matrix(~Furnishing-1, data=df) [,-1]
na_Furnishing = ifelse(data$Furnishing == "", 1, 0)
onehot_Furnishing = cbind ( onehot_Furnishing, na_Furnishing )

#i = rep(1:nrow(data))
#na_Furnishing = all(onehot_Furnishing[i,] == 0)
#1*which(onehot_Furnishing[i,] == 0)

```

## Facing

This is the cardinal direction that the home is facing. N,E,S,W, NE, NW, SE, SW.
Subjects with missing data are indicated with "1" in the last column of the one-hot matrix.

```{r mysize=T,size='\\footnotesize', echo=T,eval=T}

#summary(data$facing)
data$facing = as.factor(data$facing)
unique(data$facing)

ggplot(data, aes(x=facing)) + geom_bar()

onehot_facing = model.matrix(~facing-1, data=df) [,-1]
na_facing = ifelse(data$facing == "", 1, 0)
onehot_facing = cbind ( onehot_facing, na_facing )

```

## Overlooking

Overlooking can have multiple things listed for each home/subject. Each home may or may not overlook a "Garden/Park", "Main Road", and/or "Pool". Some homes have "Not Available" explicity listed, but others simply have no data at all for this variable.
Subjects with missing data are indicated with "1" in the last column of the one-hot matrix.

```{r mysize=T,size='\\footnotesize', echo=T,eval=T}

#summary(data$overlooking)
data$overlooking = as.factor(data$overlooking)
unique(data$overlooking)
#unique(unlist(strsplit(data$overlooking,", ")))

ggplot(data, aes(x=overlooking)) + geom_bar()

onehot_overlooking = cbind(1*grepl("Garden/Park", data$overlooking), 1*grepl("Main Road", data$overlooking), 1*grepl("Pool", data$overlooking), 1*grepl("Not Available", data$overlooking))
colnames(onehot_overlooking) = c("Garden/Park", "Main Road", "Pool", "Not Available")

na_overlooking = ifelse(data$overlooking == "", 1, 0)
onehot_overlooking = cbind ( onehot_overlooking, na_overlooking )

#ggplot(onehot_overlooking, aes(x=onehot_overlooking)) + geom_bar()

```

## Write encoded data

```{r mysize=T,size='\\footnotesize', echo=T,eval=T}

encodedData = data.frame(onehot_Furnishing, onehot_facing, onehot_overlooking)

write.csv(encodedData, "encodedData.csv")

```



